\begin{thebibliography}{1}

\bibitem{Bardram05}
J. E. Bardram. \emph{The java context awareness framework (JCAF) – a service infrastructure and programming framework for Context-Aware applications.}, 
In Pervasive Computing, pages 98–115. 2005.

\bibitem{Bourdenas10}
Themistoklis Bourdenas, Morris Sloman
\emph{Starﬁsh: Policy Driven Self-Management in Wireless Sensor Networks},
SEAMS’10, May 3–4, 2010, Cape Town, South Africa

\bibitem{Bourdenas11}
Themistoklis Bourdenas, David Wood, Petros Zerfos, Flavio Bergamaschi, Ananthram Swami, Morris Sloman
\emph{Self-adaptive Routing in Multi-hop Sensor Networks},
7th IEEE Int. Conference on Network and Service Management (CNSM 2011)

\bibitem{Costanza11}
Costanza P., Hirschfeld R. \emph{Language Constructs for Context-oriented Programming: An Overview of ContextL},
Dynamic Languages Symposium, San Diego, CA, USA, 2005.

\bibitem{Ghezzi10}
Carlo Ghezzi, Matteo Pradella, Guido Salvaneschi \emph{Programming Language Support to Context-Aware Adaptation—A Case-Study with Erlang}, 2010.

\bibitem{Hirschfeld08}
Hirschfeld R., Costanza P., Nierstrasz O. 
\emph{Context-oriented programming},
Journal of Object Technology, 2008

\bibitem{Jiang07}
Xiaofan Jiang, Jay Taneja, Jorge Ortiz, Arsalan Tavakoli, Prabal Dutta, Jaein Jeong, David Culler, Philip Levis, and Scott Shenker
\emph{An Architecture for Energy Management in Wireless Sensor Networks},
In International Workshop on Wireless Sensor Network Architecture (WSNA07), 2007

\bibitem{Kamina11}
Kamina T., Aotani T., Masuhara H. \emph{EventCJ: a context-oriented programming language with declarative event-based context transition.},
Proceedings of the 10th International Conference on Aspect-Oriented Software Development, AOSD ’11, ACM, New York, NY, USA, pp. 253–264, 2011

\bibitem{Marques11}
Lu´ıs Marques, Antonio Casimiro
\emph{Lightweight Dependable Adaptation for Wireless Sensor Networks},
Reliable Distributed Systems Workshops (SRDSW), 2011 30th IEEE Symposium on

\bibitem{Park08}
Park, P.G., Fischione, C. ; Bonivento, A. ; Johansson, K.H. ; Sangiovanni-Vincentelli, A.
\emph{Breath: a Self-Adapting Protocol for Wireless Sensor Networks in Control and Automation},
Sensor, Mesh and Ad Hoc Communications and Networks, 2008. SECON '08. 5th Annual IEEE Communications Society Conference on

\bibitem{Salvaneschi11}
Salvaneschi G., Ghezzi C., Pradella M.
\emph{Towards language-level support for self-adaptive software},
ACM Transactions on Autonomous and Adaptive Systems, 2011.

\bibitem{Salvaneschi12}
Salvaneschi G., Ghezzi C., Pradella M. 
\emph{Context-oriented  programming:  A  software  engineering  perspective},
The  Journal  of  Systems  and  Software, p.1801–1817, 2012.

\bibitem{Sehic11}
Sehic, S., Li, F., Dustdar, S. \emph{COPAL-ML: A Macro Language for Rapid Development of Context-Aware Applications in Wireless Sensor Networks}, 
SENSEA'11, 2011.

\bibitem{Steine11}
Marcel Steine, Cuong Viet Ngo, Ramon Serna Oliver, Marc Geilen, Twan Basten, Gerhard Fohler, Jean-Dominique Decotignie 
\emph{Proactive Reconﬁguration of Wireless Sensor Networks},
MSWiM’11, October 31–November 4, 2011, Miami, Florida, USA.

\bibitem{Wood08}
A. Wood, J. Stankovic, G. Virone, L. Selavo, Z. He, Q. Cao, T. Doan, Y. Wu, L. Fang, R. Stoleru 
\emph{Context-Aware Wireless Sensor Networks for Assisted-Living and Residential Monitoring}
Network, IEEE  (Volume:22 ,  Issue: 4 ), pp. 26 - 33, 2008.

\end{thebibliography}