\section{Introduction} 

Programmers design and implement Wireless Sensor Networks (WSN)
software to enable interactions in the real world at unprecedented
granularity. As such, WSN software is continuously confronted with a
range of largely unpredictable environment dynamics and changing
requirements, besides resource constraints. This demands WSN software
to \emph{adapt} to a range of different situations. Notwithstanding
the advances in WSN programming~\cite{mottola10:survey}, however,
programmers are sorely missing dedicated support to realize adaptive
WSN software.

\vspace{-6mm}\input{snippetsIntro} 

\fakepar{Example application} Consider a wildlife tracking
application~\cite{pasztor10:selective}. Sensor nodes are embedded in
collars attached to animals to study their social interactions. The
nodes are equipped with sensors to track an animal's movement, e.g.,
using GPS and accelerometers, and to detect its health conditions,
e.g., based on body temperature. Small solar panels harvest energy to
prolong a node's lifetime. A low-power short-range radio allows the
nodes to discover each other based on periodic radio beaconing.  A
node logs the radio contacts to track an animal's encounters with
other animals. The radio is also used to off-load the contact traces
when in reach of a fixed base-station.

The nodes run on batteries, making energy a precious resource that
programmers need to trade against the system's functionality,
depending on the situation. For example, sensor sampling consumes
non-negligible energy for the GPS. Depending on the desired
granularity and on the difference between consecutive GPS
readings---taken as indication of the pace of movement---programmers
need tune the GPS sampling frequency accordingly. The contact traces
can be sent directly to the base-station whenever in reach, but they
need to be stored locally otherwise.  When the battery is running low,
developers may turn the GPS sensor off to make sure the node survives
until the next encounter with a base-station, not to lose the
collected contact traces\lm{Changes to the pic required.}.

\fakepar{Problem} Taking into explicit account every possible
situation in the design and implementation of WSN software is a
challenge. Crucially, \emph{multiple combined dimensions} concurrently
determine how the software should adapt its operation, e.g., battery
levels and physical locations in our example application. Using
available approaches, this typically results in entangled
implementations that are difficult to debug, to maintain, and to
evolve.  As the number of dimensions affecting the execution (and
their combinations) grows, the implementations quickly turn into
``spaghetti code''~\cite{finne10improving}.

\putsnippet{
 caption=Example nesC implementation of adaptive functionality: \emph{several orthogonal functionality become entangled and need to share global data.},
 label=fig:nesc,
 boxname=boxnescEX
}

Fig.~\ref{fig:nesc} shows an intuitive, yet greatly simplified
example, using nesC~\cite{gay03nesc}. The code implements the behavior
needed in wildlife tracking to send contact logs to the base-station
whenever reachable, or to store them locally otherwise. Several
orthogonal concerns become intertwined and dependent on each
other. For example, determining what operating mode to
apply---implemented in line~\lstref{detect1}
to~\lstref{detect2}---rests within the same module as the actual
adaptive processing---implemented in line~\lstref{adapt1}
to~\lstref{adapt2}. Indeed, the two codes need to share global state,
in this case, the \code{base\_station\_reachable} flag. Managing such
global state rests entirely on the programmers' shoulders. Moreover,
the checks to apply before changing operating mode, such as those on
line~\lstref{check1} to~\lstref{check2}, appear interleaved with the
change of mode itself. Finally, the specific implementation of
adaptive functionality---using either \code{DataStore} or
\code{Collection}---is entirely visible from the caller module,
further coupling the two.

In such a situation, debugging, maintaining, and evolving the
implementations is going to be difficult. Modifying the code in one
place would likely require changes in several others. Alternative nesC
implementations of the functionality in Fig.~\ref{fig:nesc} are of
course possible to partly ameliorate the problem. However, qualitative
evidence gathered by looking at publicly available implementations,
e.g., within the TinyOS codebase~\cite{tinyos}, indicate that similar
implementation patterns are indeed very common.

% \fakepar{Context-oriented programming (COP)}

\fakepar{Contribution and road-map} We aim to redress this state of
affairs by enabling a notion of Context-Oriented Programming
(COP)~\cite{Hirschfeld08} in WSN software. COP fosters a strict
separation of concerns in implementing adaptive software. This is
achieved through two key notions: \emph{i)} the different situations
the software needs to operate in are mapped to different
\emph{contexts}, and \emph{ii)} the different context-dependent
behaviors are encapsulated in so called \emph{layered functions}, that
is, functions whose behavior changes---transparently to the
caller---depending on context.

COP already proved effective in creating context-aware mainstream
software, such as user interfaces~\cite{Keays03} and text
editors~\cite{Kamina11}, based on COP extensions of popular high-level
languages~\cite{Salvaneschi12}. At present, however, COP remains a far
cry from being applicable to WSNs. The resource constraints
that limit the functionality attainable with existing WSN programming
languages, for example, the inability to create run-time instances of
components, prevents applying COP in WSN programming as is.

To address this issue, we borrow concepts from COP and design
context-oriented programming abstractions for WSN software. To this
end, Section~\ref{sec:appdesign} illustrates design concepts conceived
to remain independent of a specific programming language. In doing so,
our goal is to decouple the abstractions from their concrete
realization in a language, thus facilitating their application to
multiple WSN languages. One such realization is \conesc, our own COP
extension to nesC. We choose nesC as the target language in that,
besides its widespread adoption and stable toolchain, it fosters a
node-centric view~\cite{mottola10:survey}. We argue that in most WSN
applications, adaptation decisions are most often local to individual
nodes. In illustrating \conesc, Section~\ref{sec:conesc} will
demonstrate how we render the processing in Fig.~\ref{fig:nesc} nicely
decoupled in different modules, and hence easier to debug and to
evolve.

We implement a dedicated translator, described in
Section~\ref{sec:translator}, that converts \conesc code to pure nesC.
Based on three representative applications, the results we illustrate
in Section~\ref{sec:eval} indicate that \conesc implementations are
increasingly decoupled and distinctively simpler. For example, the
analysis we perform with a model-checking tool to measure the number
of states that programmers need to deal with shows a $\approx$50\%
reduction in favor of \conesc, indicating that the latter
implementations are likely easier to debug and to maintain. Crucially,
these advantages come at a modest price: the MCU overhead when
performing calls to layered functions is negligible, while we measure
a a maximum 2.5\% (4.5\%) overhead in program (data) memory.

We conclude the paper by surveying related efforts in
Section~\ref{sec:related}, and with brief concluding remarks in
Section~\ref{sec:ending}.


% We are claiming that the notion of \emph{context} in a low-level software will
% significantly improve readability and re-usability of the code, as well as
% debugging and maintaining of the application. We also believe this notion will
% allow a developer to represent any dynamic variability. Moreover, our approach
% supports combinations of adaptation, which are usually taking a place in the
% real world.

% Consider for example a sensor network for wildlife tracking~\cite{pasztor10},
% which is intended to track animals' movements, social interactions and health
% conditions. To this end, nodes are attached to collars of animals and equipped
% with GPS and accelerometer sensors to track the movements, temperature sensor
% to monitor health conditions, and low-power short-range radio to perform a
% periodic beaconing. The latter allows the nodes to discover each other and then
% log the radio contacts to study an animal's social interactions.

% The way to create this kind of self-adaptive software is described in
% Sec.~\ref{sec:cop}, but the development would be difficult not only because of
% several adaptation dimensions, but also due to extreme resource limitations --
% the main challenge in WSNs. By addressing to these issues we introduce \conesc
% -- a context oriented extension of nesC language described in
% Sec.~\ref{sec:conesc}. To test our approach we developed a tool described in
% Sec.~\ref{sec:translator} to translate \conesc sources into nesC language. We
% discuss then our evaluation results in Sec.~\ref{sec:eval} in details.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "bare_conf"
%%% End: 
